
# Examples

```{r}
library(aorsf)
library(magrittr) # for %>%
```

`orsf()` is the entry-point of the `aorsf` package. It can be used to fit classification, regression, and survival forests. 

```{r child = "orsf-fit-intro.Rmd"}

```

## More than one way to grow a forest

You can use `orsf(no_fit = TRUE)` to make a *specification* to grow a forest instead of a fitted forest. 
```{r}
orsf_spec <- orsf(pbc_orsf, 
                  formula = time + status ~ . - id,
                  no_fit = TRUE)

orsf_spec
```

Why would you do this? Two reasons:

1. For very computational tasks, you may want to check how long it will take to fit the forest before you commit to it:

```{r}

orsf_spec %>% 
 orsf_update(n_tree = 10000) %>%
 orsf_time_to_train()

```


1. If fitting multiple forests, use the blueprint along with `orsf_train()` and `orsf_update()` to simplify your code:

```{r}

orsf_fit <- orsf_train(orsf_spec)
orsf_fit_10 <- orsf_update(orsf_fit, leaf_min_obs = 10)
orsf_fit_20 <- orsf_update(orsf_fit, leaf_min_obs = 20)

orsf_fit$leaf_min_obs
orsf_fit_10$leaf_min_obs
orsf_fit_20$leaf_min_obs

```

## tidymodels

`tidymodels` includes support for `aorsf` as a computational engine:

```{r, warning=FALSE, echo=TRUE, message=FALSE}

library(tidymodels)
library(censored)
library(yardstick)

pbc_tidy <- pbc_orsf %>% 
 mutate(event_time = Surv(time, status), .before = 1) %>% 
 select(-c(id, time, status)) %>% 
 as_tibble()

split  <- initial_split(pbc_tidy)

orsf_spec <- rand_forest() %>% 
 set_engine("aorsf") %>% 
 set_mode("censored regression")

orsf_fit <- fit(orsf_spec, 
                formula = event_time ~ ., 
                data = training(split))

```

Prediction with `aorsf` models at different times is also supported:

```{r}

time_points <- seq(500, 3000, by = 500)

test_pred <- augment(orsf_fit, 
                     new_data = testing(split), 
                     eval_time = time_points)

brier_scores <- test_pred %>% 
  brier_survival(truth = event_time, .pred)

brier_scores

roc_scores <- test_pred %>% 
  roc_auc_survival(truth = event_time, .pred)

roc_scores

```


