
Begin by fitting an ORSF ensemble:

```{r, child = 'orsf-fit-accelerated.Rmd'}

```

## Partial dependence two ways

You can compute partial dependence two ways using a random forest:

- using out-of-bag predictions for the training data

  ```{r}
  
  pd_train <- orsf_pd_oob(fit, pd_spec = list(bili = 1:5))
  
  pd_train
  
  ```

- using predictions for a new set of data

  ```{r}
  
  pd_test <- orsf_pd_new(fit, 
                         new_data = pbc_orsf_test, 
                         pd_spec = list(bili = 1:5))
  
  pd_test
  
  ```

Which should you use? Both!

- out-of-bag partial dependence indicates relationships that the model has learned during training. This is helpful if your goal is to interpret the model.

- partial dependence for new data shows how the model predicts outcomes for observations it has not seen. This is helpful if you want to test your model's reliability or fairness.

A reliable model should give roughly the same partial dependence values in its training and testing data, given the size of the training and testing data are both sufficient. 
