## Partial dependence for multiple variables

Let's re-fit our ORSF to all available data before proceeding.

```{r}

set.seed(329730)

fit <- orsf(pbc_orsf, Surv(time, status) ~ . -id)

```

partial dependence can show the expected value of a model's predictions as a function of a specific predictor, or as a function of multiple predictors. For instance, we can estimate predicted risk as a joint function of `bili`, `edema`, and `trt`:

```{r, -orsf_pd}

pred_spec = list(bili = seq(1, 5, length.out = 20),
               edema = levels(pbc_orsf_train$edema),
               trt = levels(pbc_orsf$trt))

pd_bili_edema <- orsf_pd_oob(fit, pred_spec)

library(ggplot2)

ggplot(pd_bili_edema, aes(x = bili, y = medn, col = trt, linetype = edema)) + 
 geom_line() + 
 labs(y = 'Expected predicted risk')

```

From inspection, 

- the model's predictions indicate slightly lower risk for the placebo group, and these do not seem to change much at different values of `bili` or `edema`.

- There is a clear increase in predicted risk with higher levels of `edema` and with higher levels of `bili`

- the slope of predicted risk as a function of `bili` appears highest among patients with `edema` of 0.5. Is the effect of `bili` modified by `edema` being 0.5? A quick sanity check with `coxph` suggests there is.

  ```{r}
  
  library(survival)
  
  pbc_orsf$edema_05 <- ifelse(pbc_orsf$edema == '0.5', 'yes', 'no')
  
  fit_cph <- coxph(Surv(time,status) ~ edema_05 * bili, 
                   data = pbc_orsf)
  
  anova(fit_cph)
  
  ```
  
```{r, echo = FALSE}

# in case pbc_orsf is used in downstream docs

pbc_orsf$edema_05 <- NULL

```
