
# Examples

## ANOVA importance

The default variable importance technique, ANOVA, is calculated while you fit an ORSF ensemble. 

```{r}

fit <- orsf(pbc_orsf,
                    Surv(time, status) ~ . - id)

fit

```

ANOVA is the default because it is fast, but it may not be as decisive as the permutation and negation techniques for variable selection.

## Raw VI values

the 'raw' variable importance values are stored in the fit object

```{r}

fit$importance

```

these are 'raw' because values for factors have not been aggregated into a single value. Currently there is one value for k-1 levels of a k level factor. For example, you can see edema_1 and edema_0.5 in the importance values above because edema is a factor variable with levels of 0, 0.5, and 1.

## Collapse VI across factor levels

To get aggregated values across all levels of each factor, use `orsf_vi()` with group_factors set to `TRUE` (the default)

```{r}

orsf_vi(fit)

```

## Add VI to an ORSF

You can fit an ORSF without VI, then add VI later

```{r}

fit_no_vi <- orsf(pbc_orsf,
                  Surv(time, status) ~ . - id,
                  importance = 'none')

# Note: you can't call orsf_vi_anova() on fit_no_vi because anova
# VI can only be computed while the forest is being grown.

orsf_vi_negate(fit_no_vi)

orsf_vi_permute(fit_no_vi)

```


## ORSF and VI all at once

fit an ORSF and compute vi at the same time

```{r}

fit_permute_vi <- orsf(pbc_orsf,
                        Surv(time, status) ~ . - id,
                        importance = 'permute')

# get the vi instantly (i.e., it doesn't need to be computed again)
orsf_vi_permute(fit_permute_vi)

```

You can still get negation VI from this fit, but it needs to be computed

```{r}

orsf_vi_negate(fit_permute_vi)

```

