
For classification, we fit an oblique RF to predict penguin species using `penguin` data from the magnificent `palmerpenguins` [R package](https://allisonhorst.github.io/palmerpenguins/)

```{r}
# An oblique classification RF
penguin_fit <- orsf(data = penguins_orsf,
                    n_tree = 5, 
                    formula = species ~ .)

penguin_fit

```

For regression, we use the same data but predict bill length of penguins:

```{r}
# An oblique regression RF
bill_fit <- orsf(data = penguins_orsf, 
                 n_tree = 5, 
                 formula = bill_length_mm ~ .)

bill_fit

```

My personal favorite is the oblique survival RF with accelerated Cox regression because it was the first type of oblique RF that `aorsf` provided (see [ArXiv paper](https://arxiv.org/abs/2208.01129); the paper is also published in *Journal of Computational and Graphical Statistics* but is not publicly available there). Here, we use it to predict mortality risk following diagnosis of primary biliary cirrhosis:

```{r}
# An oblique survival RF
pbc_fit <- orsf(data = pbc_orsf, 
                n_tree = 5,
                formula = Surv(time, status) ~ . - id)

pbc_fit

```
