
Begin by fitting an ORSF ensemble:

```{r, child = 'orsf-fit-accelerated.Rmd'}

```

## Three ways to compute PD and ICE

You can compute partial dependence and ICE three ways with `aorsf`: 

- using in-bag predictions for the training data

  ```{r}
  
  pd_train <- orsf_pd_inb(fit, pred_spec = list(bili = 1:5))
  
  pd_train
  
  ```

- using out-of-bag predictions for the training data

  ```{r}
  
  pd_train <- orsf_pd_oob(fit, pred_spec = list(bili = 1:5))
  
  pd_train
  
  ```

- using predictions for a new set of data

  ```{r}
  
  pd_test <- orsf_pd_new(fit, 
                         new_data = pbc_orsf_test, 
                         pred_spec = list(bili = 1:5))
  
  pd_test
  
  ```

- in-bag partial dependence indicates relationships that the model has learned during training. This is helpful if your goal is to interpret the model.

- out-of-bag partial dependence indicates relationships that the model has learned during training but using the out-of-bag data simulates application of the model to new data. if you want to test your model's reliability or fairness in new data but you don't have access to a large testing set.

- new data partial dependence shows how the model predicts outcomes for observations it has not seen. This is helpful if you want to test your model's reliability or fairness.

