```{r echo=FALSE}
library(magrittr)
```


You can compute individual conditional expectation and individual conditional expectations in three ways: 

- using in-bag predictions for the training data. In-bag individual conditional expectation indicates relationships that the model has learned during training. This is helpful if your goal is to interpret the model.

- using out-of-bag predictions for the training data. Out-of-bag individual conditional expectation indicates relationships that the model has learned during training but using the out-of-bag data simulates application of the model to new data. This is helpful if you want to test your model's reliability or fairness in new data but you don't have access to a large testing set.

- using predictions for a new set of data. New data individual conditional expectation shows how the model predicts outcomes for observations it has not seen. This is helpful if you want to test your model's reliability or fairness.

## Classification

Begin by fitting an oblique classification random forest:

```{r, child = 'orsf-fit-clsf-train-test.Rmd'}

```

Compute individual conditional expectation using out-of-bag data for `flipper_length_mm = c(190, 210)`. 

```{r}

pred_spec <- list(flipper_length_mm = c(190, 210))

ice_oob <- orsf_ice_oob(fit_clsf, pred_spec = pred_spec)

ice_oob

```

There are two identifiers in the output:

- `id_variable` is an identifier for the current value of the variable(s) that are in the data. It is redundant if you only have one variable, but helpful if there are multiple variables.

- `id_row` is an identifier for the observation in the original data.


Note that predicted probabilities are returned for each class and each observation in the data. Predicted probabilities for a given observation and given variable value sum to 1. For example, 

```{r}

ice_oob %>%
 .[flipper_length_mm == 190] %>% 
 .[id_row == 1] %>% 
 .[['pred']] %>% 
 sum()

```

## Regression

Begin by fitting an oblique regression random forest:

```{r, child = 'orsf-fit-regr-train-test.Rmd'}

```

Compute individual conditional expectation using new data for `flipper_length_mm = c(190, 210)`. 

```{r}

pred_spec <- list(flipper_length_mm = c(190, 210))

ice_new <- orsf_ice_new(fit_regr, 
                        pred_spec = pred_spec,
                        new_data = penguins_orsf_test)

ice_new

```

You can also let `pred_spec_auto` pick reasonable values like so:

```{r}

pred_spec = pred_spec_auto(species, island, body_mass_g)

ice_new <- orsf_ice_new(fit_regr, 
                        pred_spec = pred_spec,
                        new_data = penguins_orsf_test)

ice_new

```

By default, all combinations of all variables are used. However, you can also look at the variables one by one, separately, like so:

```{r}

ice_new <- orsf_ice_new(fit_regr, 
                        expand_grid = FALSE,
                        pred_spec = pred_spec,
                        new_data = penguins_orsf_test)

ice_new

```

And you can also bypass all the bells and whistles by using your own `data.frame` for a `pred_spec`. (Just make sure you request values that exist in the training data.)

```{r}

custom_pred_spec <- data.frame(species = 'Adelie', 
                               island = 'Biscoe')

ice_new <- orsf_ice_new(fit_regr, 
                        pred_spec = custom_pred_spec,
                        new_data = penguins_orsf_test)

ice_new

```


## Survival

Begin by fitting an oblique survival random forest:

```{r, child = 'orsf-fit-surv-train-test.Rmd'}

```

Compute individual conditional expectation using in-bag data for `bili = c(1,2,3,4,5)`:

```{r}
ice_train <- orsf_ice_inb(fit_surv, pred_spec = list(bili = 1:5))
ice_train
```

If you don't have specific values of a variable in mind, let `pred_spec_auto` pick for you:

```{r}
ice_train <- orsf_ice_inb(fit_surv, pred_spec_auto(bili))
ice_train
```

Specify `pred_horizon` to get individual conditional expectation at each value:

```{r}

ice_train <- orsf_ice_inb(fit_surv, pred_spec_auto(bili),
                          pred_horizon = seq(500, 3000, by = 500))
ice_train

```

Multi-prediction horizon ice comes with minimal extra computational cost. Use a fine grid of time values and  assess whether predictors have time-varying effects.
