---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

#' @srrstats {G1.2} *Project status describes current and anticipated future states of development.* 
 
```

# aorsf

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
  [![Codecov test coverage](https://codecov.io/gh/bcjaeger/aorsf/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bcjaeger/aorsf?branch=master)
[![R-CMD-check](https://github.com/bcjaeger/aorsf/workflows/R-CMD-check/badge.svg)](https://github.com/bcjaeger/aorsf/actions)
[![pkgcheck](https://github.com/bcjaeger/aorsf/workflows/pkgcheck/badge.svg)](https://github.com/bcjaeger/aorsf/actions?query=workflow%3Apkgcheck)
[![Status at rOpenSci Software Peer Review](https://badges.ropensci.org/532_status.svg)](https://github.com/ropensci/software-review/issues/532)
<!-- badges: end -->

`aorsf` provides optimized software to fit, interpret, and make predictions with oblique random survival forests (ORSFs).  


## Why aorsf?

- over 400 times faster than `obliqueRSF`.

- accurate predictions for time-to-event outcomes.

- negation importance, a novel technique to estimate variable importance for ORSFs.

- intuitive API with formula based interface.

- extensive input checks + informative error messages.

## Installation

You can install the development version of aorsf from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("bcjaeger/aorsf")
```
## Example

The `orsf()` function is used to fit ORSFs. Printing the output from `orsf()` will give some descriptive statistics about the ensemble.

```{r example}

library(aorsf)

fit <- orsf(data_train = pbc_orsf,
            formula = Surv(time, status) ~ . - id)

print(fit)

```

How about interpreting the fit? 

- use `orsf_vi_negate()` and `orsf_vi_anova()` for variable importance

  ```{r}
  
  orsf_vi_negate(fit)
  
  ```

- use `orsf_pd_ice()` or `orsf_pd_summary()` for individual or aggregated partial dependence values. 

  ```{r}
  
  orsf_pd_summary(fit, pd_spec = list(bili = c(1:5)))
  
  ```


- use `orsf_summarize_uni()` to show the top predictor variables in an ORSF model and the expected predicted risk at specific values of those predictors. (The term 'uni' is short for univariate.)

  ```{r}

  # take a look at the top 5 variables 
  # for continuous predictors, see expected risk at 25/50/75th quantile
  # for categorical predictors, see expected risk in specified category

  orsf_summarize_uni(object = fit, n_variables = 5)

  ```

## References

Byron C. Jaeger, D. Leann Long, Dustin M. Long, Mario Sims, Jeff M. Szychowski, Yuan-I Min, Leslie A. Mcclure, George Howard, Noah Simon (2019). Oblique Random Survival Forests. Ann. Appl. Stat. 13(3): 1847-1883. URL https://doi.org/10.1214/19-AOAS1261 DOI: 10.1214/19-AOAS1261

## Funding

The developers of `aorsf` receive financial support from the Center for Biomedical Informatics, Wake Forest University School of Medicine. We also receive support from the National Center for Advancing Translational Sciences of the National Institutes of Health under Award Number UL1TR001420. 

The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health.

