---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aorsf

<!-- badges: start -->
  [![Codecov test coverage](https://codecov.io/gh/bcjaeger/aorsf/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bcjaeger/aorsf?branch=master)
[![R-CMD-check](https://github.com/bcjaeger/aorsf/workflows/R-CMD-check/badge.svg)](https://github.com/bcjaeger/aorsf/actions)
<!-- badges: end -->

The goal of `aorsf` is to fit, interpret, and make predictions with oblique random survival forests (ORSFs).

## Installation

You can install the development version of aorsf from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("bcjaeger/aorsf")
```
## Example

The `orsf()` function is used to fit ORSFs. Printing the output from `orsf()` will give some descriptive statistics about the ensemble.

```{r example}
library(aorsf)

fit <- orsf(data_train = pbc_orsf,
            formula = Surv(time, status) ~ . - id)

print(fit)

```

How about interpreting the fit? There are several functions for this: `orsf_vi()` for variable importance, `orsf_interaction()` for two-way variable interactions, and `orsf_pd_ice()` or `orsf_pd_summary()` for individual or aggregated partial dependence values. However, `orsf_summarize_uni()` is the most convenient way to assess top predictor variables in ORSF and the expected predicted risk at specific values of those predictors.

```{r}

# take a look at the top 5 variables 
# for continuous predictors, see expected risk at 25/50/75th quantile
# for categorical predictors, see expected risk in each category

orsf_summarize_uni(object = fit, n_variables = 5)

```

The term 'uni' is short for univariate.

