---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aorsf

<!-- badges: start -->
  [![Codecov test coverage](https://codecov.io/gh/bcjaeger/aorsf/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bcjaeger/aorsf?branch=master)
[![R-CMD-check](https://github.com/bcjaeger/aorsf/workflows/R-CMD-check/badge.svg)](https://github.com/bcjaeger/aorsf/actions)
<!-- badges: end -->

The goal of `aorsf` is to fit, interpret, and make predictions with oblique random survival forests (ORSFs). The 'a' in the title stands for accelerated. So why do ORSFs need to be accelerated? Oblique decision trees are often more accurate but slower to fit compared to their axis-based counterparts. The issue of higher computation time is compounded for survival decision trees, which usually require more computing than classification or regression trees. To make ORSF more accessible and able to engage with larger datasets, `aorsf` applies strategies to cut down computing time without sacrificing prediction accuracy.  

## Installation

You can install the development version of aorsf from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("bcjaeger/aorsf")
```
## Example

The `orsf()` function is used to fit ORSFs. Printing the output from `orsf()` will give some descriptive statistics about the ensemble.

```{r example}
library(aorsf)

fit <- orsf(data_train = pbc_orsf,
            formula = Surv(time, status) ~ . - id)

print(fit)

```

How about interpreting the fit? There are several functions for this: `orsf_vi()` for variable importance, `orsf_interaction()` for two-way variable interactions, and `orsf_pd_ice()` or `orsf_pd_summary()` for individual or aggregated partial dependence values. However, `orsf_summarize_uni()` is the most convenient way to assess top predictor variables in ORSF and the expected predicted risk at specific values of those predictors. The term 'uni' is short for univariate.

```{r}

# take a look at the top 5 variables 
# for continuous predictors, see expected risk at 25/50/75th quantile
# for categorical predictors, see expected risk in each category

orsf_summarize_uni(object = fit, n_variables = 5)

```

## References

Byron C. Jaeger, D. Leann Long, Dustin M. Long, Mario Sims, Jeff M. Szychowski, Yuan-I Min, Leslie A. Mcclure, George Howard, Noah Simon (2019). Oblique Random Survival Forests. Ann. Appl. Stat. 13(3): 1847-1883. URL https://doi.org/10.1214/19-AOAS1261 DOI: 10.1214/19-AOAS1261
