---
title: "Introduction to aorsf"
description: >
  Learn how to get started with the basics of aorsf.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to aorsf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5, 
  fig.width = 7
)
```

This article covers core features of the `aorsf` package. Separate articles will cover these features in more detail. 

## Background: ORSF

The oblique random survival forest (ORSF) is an extension of the axis-based RSF algorithm. Axis based RSFs can be fit with several R packages, including `RandomForestSRC`, `party`, and `ranger`. However, apart from `aorsf` and `obliqueRSF`, few R packages fit ORSFs.


## Accelerated ORSF

The purpose of `aorsf` ('a' is short for accelerated) is to provide routines to fit ORSFs that will scale adequately to large data sets. For example, `aorsf::orsf()` runs hundreds of times faster than its predecessor, `obliqueRSF::ORSF()`.

To fit an accelerated ORSF model, use the `orsf` function:

```{r}

library(aorsf)
library(ggplot2)

set.seed(329)

orsf_fit <- orsf(data = pbc_orsf, 
                 formula = Surv(time, status) ~ . - id, 
                 n_tree = 2500)

orsf_fit

```

you may notice that the first input of `aorsf` is `data`. This is a design choice that makes it easier to use `orsf` with pipes (i.e., `%>%` or `|>`). For instance,

```{r, eval=FALSE}
library(dplyr)

orsf_fit <- pbc_orsf |> 
 select(-id) |> 
 orsf(formula = Surv(time, status) ~ .)

```

## Interpretation

`aorsf` includes several functions dedicated to interpretation of ORSFs, both through estimation of partial dependence and variable importance.

### Variable importance

`aorsf` provides multiple ways to compute variable importance. 

- To compute negation importance, ORSF multiplies each coefficient of that variable by -1 and then re-computes the out-of-sample (sometimes referred to as out-of-bag) accuracy of the ORSF model. 

    ```{r}

    orsf_vi_negate(orsf_fit)

    ```

- You can also compute variable importance using permutation, a more classical approach.

    ```{r}
  
    orsf_vi_permute(orsf_fit)
  
    ```

- A faster alternative to permutation and negation importance is ANOVA importance, which computes the proportion of times each variable obtains a low p-value (p < 0.01) while the forest is grown.

    ```{r}
    
    orsf_vi_anova(orsf_fit)
    
    ```


### Summarize ORSF models

Getting straight to variable importance and partial dependence in a direct way can often be helpful in exploratory analyses. When you want a lot of information for just a little code, `orsf_summarize_uni()` is the tool for you:

```{r}
orsf_summarize_uni(orsf_fit, 
                   n_variables = 5, 
                   pred_horizon = 365.25 * 5)
```

## What about the original ORSF?

The original ORSF (i.e., `obliqueRSF`) used `glmnet` to find linear combinations of inputs. `aorsf` allows users to implement this approach using the `orsf_control_net()` function: 

```{r}

orsf_net <- orsf(data = pbc_orsf, 
                 formula = Surv(time, status) ~ . - id, 
                 control = orsf_control_net(),
                 n_tree = 50)

```

`net` forests fit a lot faster than the original ORSF function in `obliqueRSF`. However, `net` forests are still much slower than `cph` ones:

```{r}

# tracking how long it takes to fit 50 glmnet trees
print(
 t1 <- system.time(
  orsf(data = pbc_orsf, 
       formula = Surv(time, status) ~ . - id, 
       control = orsf_control_net(),
       n_tree = 50)
 )
)

# and how long it takes to fit 50 cph trees
print(
 t2 <- system.time(
  orsf(data = pbc_orsf, 
       formula = Surv(time, status) ~ . - id, 
       control = orsf_control_cph(),
       n_tree = 50)
 )
)

t1['elapsed'] / t2['elapsed']


```

## aorsf and other machine learning software

Generally, `aorsf` follows conventions set by the R packages `randomForest` and `randomForestSRC`. For example, the `print` method for `aorsf` models is very similar to `randomForestSRC` models, as is the specification of computing variable importance and the objects returned by variable importance functions. Similar to `randomForestSRC`, `aorsf` allows and recommends partial dependence to be estimated using out-of-bag data. 

The unique feature of `aorsf` is fitting oblique random survival forests fast while maintaining strong prediction accuracy. `RLT` and `obliqueRSF` both fit oblique random survival forests, but `aorsf` does so faster. `ranger` and `randomForestSRC` fit survival forests, but neither package supports oblique splitting. `obliqueRF` fits oblique random forests for classification and regression, but not survival. `PPforest` fits oblique random forests for classification but not survival. The default prediction behavior for `aorsf` models is to produce predicted risk at a specific prediction horizon, which is not the default for `ranger`  or `randomForestSRC`.

