---
title: "Tips to speed up computation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tips to speed up computation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(aorsf)
```

## Go faster

Analyses can slow to a crawl when models need hours to run. In this article you will find a few tricks to prevent this bottleneck when using `orsf()`.

## Use `orsf_control_fast()`

This is the default `control` value for `orsf()` and its run-time compared to other approaches can be striking. For example:


```{r}

time_fast <- system.time(
 expr = orsf(pbc_orsf, 
             formula = time+status~. -id, 
             control = orsf_control_fast(), 
             n_tree = 5)
)

time_net <- system.time(
 expr = orsf(pbc_orsf, 
             formula = time+status~. -id, 
             control = orsf_control_net(), 
             n_tree = 5)
)

# control_fast() is much faster
time_net['elapsed'] / time_fast['elapsed']

```

## Use `n_thread`

The `n_thread` argument uses multi-threading to run `aorsf` functions in parallel when possible. If you know how many threads you want, e.g. you want exactly 5, just say `n_thread = 5`. If you aren't sure how many threads you have available but want to use as many as you can, say `n_thread = 0` and `aorsf` will figure out the number for you.

```{r}

# automatically pick number of threads based on amount available

orsf(pbc_orsf, 
     formula = time+status~. -id, 
     n_tree = 5,
     n_thread = 0)

```

Because R is a single threaded language, multi-threading cannot be applied when `orsf()` needs to call R functions from C++, which occurs when a customized R function is used to find linear combination of variables or compute prediction accuracy.

## Do less

There are some defaults in `orsf()` that can be adjusted to make it run faster:

- set `n_retry` to 0 instead of 3 (the default)

- set `oobag_pred_type` to 'none' instead of 'surv' (the default)

- set 'importance' to 'none' instead of 'anova' (the default)

- increase `split_min_events`, `split_min_obs`, `leaf_min_events`, or `leaf_min_obs` to make trees stop growing sooner

- increase `split_min_stat` to make trees stop growing sooner

Applying these tips:

```{r}

orsf(pbc_orsf, 
     formula = time+status~., 
     na_action = 'na_impute_meanmode',
     n_thread = 0, 
     n_tree = 5, 
     n_retry = 0,
     oobag_pred_type = 'none', 
     importance = 'none',
     split_min_events = 20, 
     leaf_min_events = 10,
     split_min_stat = 10)

```

While these default values do make `orsf()` run slower, they also usually make its predictions more accurate or make the fit easier to interpret. 

## Show progress

Setting `verbose_progress = TRUE` doesn't make anything run faster, but it can help make it *feel* like things are running less slow.

```{r}

verbose_fit <- orsf(pbc_orsf, 
                    formula = time+status~. -id, 
                    n_tree = 5, 
                    verbose_progress = TRUE)

```


