---
title: "Partial dependence and ICE curves with ORSF"
description: >
  Learn how to interpret ORSF models using partial dependence.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Partial dependence and ICE curves with ORSF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.height = 5, 
  fig.width = 7
)

library(aorsf)

docs_pd <- c("D:/aorsf/Rmd/orsf_pd_content_1.Rmd",
             "D:/aorsf/Rmd/orsf_pd_content_2.Rmd")

docs_ice <- c("D:/aorsf/Rmd/orsf_ice_content_1.Rmd",
              "D:/aorsf/Rmd/orsf_ice_content_2.Rmd")

```

## Partial dependence

`r aorsf:::roxy_pd_explain()`

Begin by fitting an ORSF ensemble.

```{r}

library(aorsf)

set.seed(329730)

index_train <- sample(nrow(pbc_orsf), 150) 

pbc_orsf_train <- pbc_orsf[index_train, ]
pbc_orsf_test <- pbc_orsf[-index_train, ]

fit <- orsf(data = pbc_orsf_train, 
            formula = Surv(time, status) ~ . - id,
            oobag_pred_horizon = 365.25 * 5)

fit

```

## Partial dependence two ways

You can compute partial dependence two ways using a random forest:

- using out-of-bag predictions for the training data

  ```{r}
  
  pd_train <- orsf_pd_oob(fit, pd_spec = list(bili = 1:5))
  
  pd_train
  
  ```

- using predictions for a new set of data

  ```{r}
  
  pd_test <- orsf_pd_new(fit, 
                         new_data = pbc_orsf_test, 
                         pd_spec = list(bili = 1:5))
  
  pd_test
  
  ```

Which should you use? Both!

- out-of-bag partial dependence indicates relationships that the model has learned during training. This is helpful if your goal is to interpret the model.

- partial dependence for new data shows how the model predicts outcomes for observations it has not seen. This is helpful if you want to test your model's reliability or fairness.

A reliable model should give roughly the same partial dependence values in its training and testing data, given the size of the training and testing data are both sufficient. 

## Partial dependence for multiple variables

Let's re-fit our ORSF to all available data before proceeding.

```{r}

set.seed(329730)

fit <- orsf(pbc_orsf, Surv(time, status) ~ . -id)

```


partial dependence can show the expected value of a model's predictions as a function of a specific predictor, or as a function of multiple predictors. For instance, we can estimate predicted risk as a joint function of `bili`, `edema`, and `trt`:

```{r, -orsf_pd}

pd_spec = list(bili = seq(1, 5, length.out = 20),
               edema = levels(pbc_orsf_train$edema),
               trt = levels(pbc_orsf$trt))

pd_bili_edema <- orsf_pd_oob(fit, pd_spec)

library(ggplot2)

ggplot(pd_bili_edema, aes(x = bili, y = medn, col = trt, linetype = edema)) + 
 geom_line() + 
 labs(y = 'Expected predicted risk')

```

From inspection, 

- the model's predictions indicate slightly lower risk for the placebo group, and these do not seem to change much at different values of `bili` or `edema`.

- There is a clear increase in predicted risk with higher levels of `edema` and with higher levels of `bili`

- the slope of predicted risk as a function of `bili` appears highest among patients with `edema` of 0.5. Is the effect of `bili` modified by `edema` being 0.5? A quick sanity check with `coxph` suggests there is.

  ```{r}
  
  library(survival)
  
  pbc_orsf$edema_05 <- ifelse(pbc_orsf$edema == '0.5', 'yes', 'no')
  
  fit_cph <- coxph(Surv(time,status) ~ edema_05 * bili, 
                   data = pbc_orsf)
  
  anova(fit_cph)
  
  ```
  
```{r, echo = FALSE}

# in case pbc_orsf is used in downstream docs

pbc_orsf$edema_05 <- NULL

```



## Individual conditional expectations

`r aorsf:::roxy_ice_explain()`



