---
title: "Compute partial dependence with ORSF"
description: >
  Learn how to interpret ORSF models using partial dependence.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compute partial dependence with ORSF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.height = 5, 
  fig.width = 7
)
```

```{r setup}
library(aorsf)
library(ggplot2)
```

This article is under construction

## Background: Partial dependence

Partial dependence plots show the average marginal effect a combination of features imparts on predicted risk. Individual conditional expectation plots narrow the focus to individual datapoints. 

## Getting partial dependance from an `aorsf` object

First fit an accelerated ORSF model on `pbc_orsf`, a modification of the Mayo Clinic Primary Biliary Cholangitis data:

```{r}

set.seed(123)

orsf_fit <- orsf(data_train = pbc_orsf,
                 formula = Surv(time, status) ~ . -id,
                 n_tree = 2500)

```

To view the partial dependence between predicted risk and the combination of blood vessel malformations in the skin (`spiders`) and urine copper (`copper`), we use `orsf_pd_summary`. The `pd_spec` argument contains the variable values we want to compute partial dependence for. By default `oobag == TRUE`, meaning partial dependence is calculated using the out of bag training data. We keep this set to `TRUE` because we are computing partial dependence using the training data. `expand_grid == TRUE` tells `orsf_pd_summary` to use all combinations of values in `pd_spec`.

```{r}
pd_spec <- list(spiders = c("0", "1"),
                copper = seq(45, 130, by = 10))

pd_data <- orsf_pd_summary(object = orsf_fit,
                           pd_spec = pd_spec,
                           oobag = TRUE,
                           expand_grid = TRUE)

ggplot(pd_data) +
 aes(x = copper, y = mean, col = spiders) +
 geom_line() + 
 labs(y = 'Predicted risk',
      x = 'Urine Copper',
      title = 'Partial dependence of copper and spiders')
```

Next lets look at Individual Conditional Expectation (ICE). Each instance in the dataset is represented by one line showing change in prediction dependent on change in a feature. While `orsf_pd_ice` allows you to use multiple features, it is recommended to view one at a time. In the plot below we can see how each data point's predicted survival changes as copper changes. 


```{r}

ice_data <- orsf_pd_ice(object = orsf_fit,
                       pd_spec = list(copper = seq(45, 130, by = 10)),
                       expand_grid = TRUE)

ice_data[id_row %in% c(1:30)] |> 
 ggplot(aes(x=copper, y=pred, group=id_row)) + 
 geom_line() +
 labs(y = 'Predicted risk',
     x = 'Urine Copper',
     title = 'Individual conditional expectation of copper',
     subtitle = 'Data for 30 patients are shown')
 
```

